Robots Showing Emotions: Emotion Representation with no bio-inspired Body

### 2016-10-30 JournalInteractionStudies/Rebuttal2.txt

General
* IT was corrected Cañamero reference
* It was added the following article as reference: Relationship between color and emotion
* Introduction parapragh on Case study section was extended to include the conexion and create a better connection among the papers.

* One sentece to introduce next case study was added at the end of each case study

Reviewer 1
*****************************************

>>Q: First case study: There is no discussion of the results. Whilst the statistical analysis is now clearer, the authors should explain what has been learned in relation to theses results. Why were Happiness and Anger recognized? Do the authors have some hypothesis? Can any generalization be drawn?

>>Q: Second case study: What were the scenarios?
A: a link to a video, which shows the emotions, was added
>>Q: "However, if the conveyed emotion does not match the scene, the audience does not recognize it.": How can the design allow for such claim? Is this supported by the results? Is it possible that only the context was recognized regardless of what the robot was doing?
A: the sentence were changed, to "this opens the question about the influence that could have the human actor and the scene on the recognition rate"
>>Q: Third one: Why was the questionnaire changed half way?
It is well established that velocity is related to arousal. Why additional features are needed? The literature suggests that it is enough to discriminate between low and high velocity.
A: The re
Reviewer 3
*****************************************
Case study titles were changed:
First Case study: Empirical Emotion Parameters
Second Case Study: Addition of Context
Third Case Study: Literature Emotional Parameters

### 2016-10-30 JournalInteractionStudies/Journal-AISB.tex
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.
Changes in some parts of the article, main changes are described in Rebuttal2.txt

\abstract{Robots should be able to represent emotional states to interact with people as social agents. There are cases where robots cannot have bio-inspired bodies, for instance because the task to be performed requires a special shape, as in the case of home cleaners, package carriers, and many others. In these cases, emotional states have to be represented by exploiting movements of the body. In this paper, we present a set of case studies aimed to identify specific values to convey emotion though changes in linear and angular velocities, which might be applied on different non-bio-inspired bodies. The work originates from some of the most considered emotion expression theories and from emotion coding for people. We show that people can recognize some emotional expressions better than others, and we propose some directions to express emotions exploiting only bio-neutral movement.
}

>>>>

%TODO (Done)change non-human like to non-antrhropomorphic
%TODO (Done)add a paragrahp at begining of section Case studies explaining the path followed
%TODO (Done) correct the paragraph written
%TODO (Done) make modification to case studies subtitles to give more information
%TODO (Done)review reference: . A. Russell and J.-A. Bachorowski et al., "Facial and vocal expressions of emotion," Annu. Rev. Psychol., vol. 54, no. 1, pp. 329-349, 2003.
%TODO Write a paragraph with this information
%TODO (Done)Add the relevant case study number to Table 2, 4, 6, 7, 11
%TODO Pg. 18, line 9. Add the word "respectively" at the end, since it could also be interpreted as "low and high arousal emotions are difficult to discriminate between each other".
%TODO (Done) correct Cañamero reference
%TODO (Done) find a reference for color 
%TODO (Done)make all the names for emotions expressed the same way (e.g. capital initial and italic, or only capital initial, or only italic, or nothing).

\abstract{Robots should be able to represent emotional states to interact with people as social agents. There are cases where robots cannot have bio-inspired bodies, for instance because the task to be performed requires a special shape, as in the case of home cleaners, package carriers, and many others. In these cases, emotional states have to be represented by exploiting movements of the body. In this paper, we present a set of case studies aimed at identifying specific values to convey emotion trough changes in linear and angular velocities, which might be applied on different non-anthropomorphic bodies. This work originates from some of the most considered emotion expression theories and from emotion coding for people. We show that people can recognize some emotional expressions better than others, and we propose some directions to express emotions exploiting only bio-neutral movement.
}

### 2016-10-30 JournalInteractionStudies/Parts/Emotion_Projection.tex
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.
Changes in some parts of the article, main changes are described in Rebuttal2.txt

Studies focused on robotics bodies can be partitioned in two groups: platforms that resemble humans or animals, and those that don't. Moreover, the implemented emotions vary among works, with a special tendency to include happiness. Questionnaires to assess people's perception are often adopted in works that use human or animal like platforms~\cite{Canamero2010,Beck2010,Li2011,Destephe2013b,Arras2012,Brown2014}. A relevant exception is the work done by Lakatos and collaborators~\cite{Lakatos2014} who assessed the participant's perception based on interaction through a play. In the experiment, each participant had two balls (yellow and black) and he could play with the robot using one of this balls. However, one of the balls would trigger robot's happiness and the other sadness.

Regarding the movement or poses design, three major approaches could be found. In the first approach, actors are asked to perform different walks conveying emotions~\cite{Destephe2013b}. These movements are later reproduced with the robot, as closely as possible. A second approach is to ask some subjects to design the movements~\cite{Li2011}. The last approach is the empirical approach, where researches come with poses and movements based on their own experience~\cite{Canamero2010,Beck2010,Arras2012,Brown2014}. One major finding from diverse works~\cite{Canamero2010,Beck2010,Brown2014} is that moving the head up improved the identification of pride, happiness, and excitement. While moving the head down improved the identification of anger and sadness.

On the other hand, the works that use platforms with no resemblance to persons or animals could be sub-divided in three groups. A first group study the possibility to convey emotions with their platforms~\cite{Arras2012, Novika2015, BarakovaL10}. The second group focus on determining the contribution of different features in the perception of emotions~\cite{Saerbeck2010,Barakova2013, Sharma2013, NAM2014}.
The second group use pre-defined procedures to assess the participants' perception, such as Self-Assessment Manikin (SAM)~\cite{Lang2008}, possibly combined  with other procedures such as Social Dominance Orientation (SDO)~\cite{pratto1994social} and PANAS~\cite{WatsonClarkTellegen88}.

It is important to mention the use of Laban’s Efforts~\cite{Laban1968} in the selection of the parameters in few works~\cite{BarakovaL10, Sharma2013}. Besides the fact that these works found that it was possible to convey emotions with non-human- or animal-like platforms, a relevant finding was the identification of the correlation between acceleration and speed with the arousal axis in the circumplex model of affect~\cite{Petta2010}.

Table~\ref{table:comparison_work_emotion_projection} summarizes all the robotics studies already mentioned, highlighting the following characteristics:
\begin{itemize}
	\item \textit{Embodiment} tells if the robot is human-like, robot-like or neither of these two;
	\item \textit{Platform used} gives the name of the platform used on the study;
	\item The\textit{Locomotion} gives information about the capabilities of the platform., because, even having a human-like embodiment, it does not mean that the platform can move;
	\item \textit{Emotion Evaluation Method} refers to the methodology used to collect the data in the studies.
	\item \textit{Emotion presentation} tells if a real robot was used in the study, or any other method;
	\item \textit{Emotions Implemented} gives the list of emotions showed to the subjects;
	\item \textit{How is the emotion conveyed?} This tells what medium was used to convey the emotion. The mediums considered were: movement, body posture and face poses.
\end{itemize}
\clearpage
\begin{table}[h]
\caption{Comparison among works on emotion projection in robotics. NA = Not Available}
\label{table:comparison_work_emotion_projection}
\begin{center}
\tiny %\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\begin{tabular}{|p{2.3 cm}|p{1.5 cm}|p{1.3 cm}|p{1.4 cm}|p{1.4 cm}|p{1.0 cm}|p{2.2 cm}|p{1.4 cm}|}
%\begin{tabular}{|p{1.5 cm}|p{1.0 cm}|p{1.5 cm}|p{1.5 cm}|p{1.5 cm}|p{2.0 cm}|p{1.5}|p{2.0 cm}|p{1.0 cm}|p{1.5 cm}|}
\hline 
\textbf{Work}  & \textbf{Embodiment} & \textbf{Platform Used} & \textbf{Locomotion} & \textbf{Emotion Evaluation Method} & \textbf{Emotion Presentation} & \textbf{Emotions Implemented}  & \textbf{How is the emotion conveyed?}\\ 
\hline 
Breazeal~\cite{Breazeal2002}  & Face & Kismet & NA & Questionnaire & Video & Happiness, surprise, anger, disgust, fear, sadness and interest &  Face poses\\ 
\hline
Saerbeck and Christoph~\cite{Saerbeck2010} & Animal / Non-Human like& iCat / Roomba & NA / Differential &PANAS and SAM & Real Robot & NA &  Movement\\
\hline 
Ca\~namero and collaborators~\cite{Canamero2010,Beck2010}&Human-Like & NAO & Bipedal & Questionnaire & Real robot & Anger, sadness, fear, pride, happiness, and excitement &  Body poses\\
\hline
Li and Chignell~\cite{Li2011}& Animal-Like& Teddy Bear Robot& NA & Questionnaire & Video (Real robot) & Random + Anger, disgust, fear, happiness, sadness, and surprise  & Body poses\\ 
\hline
Barakova and collaborators~\cite{Barakova2013} & NA & Closet robot& NA & SDO and SAM & Real robot & NA*  & Changing lights on and intensity\\ 
\hline
Sharma and collaborators~\cite{Sharma2013} & Non-Human / Animal & Quadrotor & Aerial & SAM + Questionnaire + Interview &  Real Robot & Laban's poles & Movement\\
\hline
Destephe and collaborators~\cite{Destephe2013b}&Human-like & WABIAN-2R & Bipedal & Questionnaire & Video (Virtual Robot) & Fear, anger, happiness, and sadness & Movement (Gait)+Body Poses\\
\hline 
Embgen and collaborators~\cite{Arras2012} & Human-like & Daryl &  Differential& Questionnaire & Real Robot & Happiness, sadness, fear, curiosity, embarrassment, and disappointment & Movement + Body poses \\ 
\hline
Lakatos and Collaborators~\cite{Lakatos2014}& Animal-Like & Dog & Holonomic & Indirect (Interaction with the robot) & Real Robot & Happiness and Fear &  Move + Body poses \\
\hline
Brown and Howard~\cite{Brown2014}& Human-Like & DARwIn-OP & Bipedal & Questionnaire & Real Robot & Happiness and sadness & Body poses\\
\hline
Novikova and Watss~\cite{Novika2015}&Non-Human/ Animal & Undetermine & Differential& Questionnaire& Real Robot & Scared, surprise, excited, angry, neutral, happiness and sadness &  Movement + Poses\\
\hline 
\end{tabular} 
\end{center}
\end{table}

>>>>

Studies focused on robotics bodies can be partitioned in two groups: platforms that resemble humans or animals, and those that don't. Moreover, the implemented emotions vary among works, with a tendency to include happiness. Questionnaires to assess people's perception are often adopted in works that use anthropomorphic platforms~\cite{Canamero2010,Beck2010,Li2011,Destephe2013b,Arras2012,Brown2014}. A relevant exception is the work done by Lakatos and collaborators~\cite{Lakatos2014} who assessed the participant's perception based on interaction through a play. In the experiment, each participant had two balls (yellow and black) and he could play with the robot using one of this balls. One of the balls would trigger robot's happiness and the other sadness.

Regarding the movement or pose design, three major approaches could be found. In the first approach, actors are asked to perform different walks conveying emotions~\cite{Destephe2013b}. These movements are later reproduced with the robot, as closely as possible. A second approach is to ask some subjects to design the movements~\cite{Li2011}. The last approach is the empirical approach, where researches come with poses and movements based on their own experience~\cite{Canamero2010,Beck2010,Arras2012,Brown2014}. One major finding from diverse works~\cite{Canamero2010,Beck2010,Brown2014} is that moving the head up improved the identification of pride, happiness, and excitement. While moving the head down improved the identification of anger and sadness.

On the other hand, the works that use platforms with no resemblance to persons or animals could be sub-divided in three groups. A first group study the possibility to convey emotions with their platforms~\cite{Arras2012, Novika2015, BarakovaL10}. The second group focus on determining the contribution of different features in the perception of emotions~\cite{Saerbeck2010,Barakova2013, Sharma2013, NAM2014}.
This group use pre-defined procedures to assess the participants' perception, such as Self-Assessment Manikin (SAM)~\cite{Lang2008}, possibly combined  with other procedures such as Social Dominance Orientation (SDO)~\cite{pratto1994social} and PANAS~\cite{WatsonClarkTellegen88}.

It is important to mention the use of Laban’s Efforts~\cite{Laban1968} in the selection of the parameters in few works~\cite{BarakovaL10, Sharma2013}. Besides the fact that these works report that it was possible to convey emotions with non-human- or animal-like platforms, a relevant finding was the identification of the relationship between acceleration and speed with the arousal axis in the circumplex model of affect~\cite{Petta2010}.

Table~\ref{table:comparison_work_emotion_projection} summarizes all the robotics studies already mentioned, highlighting the following characteristics:
\begin{itemize}

	\item \textit{Embodiment} tells if the robot is human-like, robot-like or neither of these two;

	\item \textit{Platform used} gives the name of the platform used on the study;

	\item The\textit{Locomotion} gives information about the capabilities of the platform., because, even having a human-like embodiment, it does not mean that the platform can move;

	\item \textit{Emotion Evaluation Method} refers to the methodology used to collect the data in the studies.

	\item \textit{Emotion presentation} tells if a real robot was used in the study, or any other method;

	\item \textit{Emotions Implemented} gives the list of emotions showed to the subjects;

	\item \textit{How is the emotion conveyed?} This tells what medium was used to convey the emotion. The mediums considered were: movement, body posture and face poses.

\end{itemize}

\clearpage
\begin{table}[h]
\caption{Comparison among works on emotion projection in robotics. NA = Not Available}
\label{table:comparison_work_emotion_projection}
\begin{center}
\tiny %\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\begin{tabular}{|p{2.3 cm}|p{1.5 cm}|p{1.3 cm}|p{1.4 cm}|p{1.4 cm}|p{1.0 cm}|p{2.2 cm}|p{1.4 cm}|}
%\begin{tabular}{|p{1.5 cm}|p{1.0 cm}|p{1.5 cm}|p{1.5 cm}|p{1.5 cm}|p{2.0 cm}|p{1.5}|p{2.0 cm}|p{1.0 cm}|p{1.5 cm}|}
\hline 
\textbf{Work}  & \textbf{Embodiment} & \textbf{Platform Used} & \textbf{Locomotion} & \textbf{Emotion Evaluation Method} & \textbf{Emotion Presentation} & \textbf{Emotions Implemented}  & \textbf{How is the emotion conveyed?}\\ 
\hline 
Breazeal~\cite{Breazeal2002}  & Face & Kismet & NA & Questionnaire & Video & Happiness, surprise, anger, disgust, fear, sadness and interest &  Face poses\\ 
\hline
Saerbeck and Christoph~\cite{Saerbeck2010} & Animal / Non-Human like& iCat / Roomba & NA / Differential &PANAS and SAM & Real Robot & NA &  Movement\\
\hline 
Ca\~namero and collaborators~\cite{Canamero2010,Beck2010}&Human-Like & NAO & Bipedal & Questionnaire & Real robot & Anger, sadness, fear, pride, happiness, and excitement &  Body poses\\
\hline
Li and Chignell~\cite{Li2011}& Animal-Like& Teddy Bear Robot& NA & Questionnaire & Video (Real robot) & Random + Anger, disgust, fear, happiness, sadness, and surprise  & Body poses\\ 
\hline
Barakova and collaborators~\cite{Barakova2013} & NA & Closet robot& NA & SDO and SAM & Real robot & NA*  & Changing lights on and intensity\\ 
\hline
Sharma and collaborators~\cite{Sharma2013} & Non-Human / Animal & Quadrotor & Aerial & SAM + Questionnaire + Interview &  Real Robot & Laban's poles & Movement\\
\hline
Destephe and collaborators~\cite{Destephe2013b}&Human-like & WABIAN-2R & Bipedal & Questionnaire & Video (Virtual Robot) & Fear, anger, happiness, and sadness & Movement (Gait)+Body Poses\\
\hline 
Embgen and collaborators~\cite{Arras2012} & Human-like & Daryl &  Differential& Questionnaire & Real Robot & Happiness, sadness, fear, curiosity, embarrassment, and disappointment & Movement + Body poses \\ 
\hline
Lakatos and Collaborators~\cite{Lakatos2014}& Animal-Like & Dog & Holonomic & Indirect (Interaction with the robot) & Real Robot & Happiness and Fear &  Move + Body poses \\
\hline
Brown and Howard~\cite{Brown2014}& Human-Like & DARwIn-OP & Bipedal & Questionnaire & Real Robot & Happiness and sadness & Body poses\\
\hline
Novikova and Watss~\cite{Novika2015}&Non-Human/ Animal & Undetermine & Differential& Questionnaire& Real Robot & Scared, surprise, excited, angry, neutral, happiness and sadness &  Movement + Poses\\
\hline 
\end{tabular} 
\end{center}
\end{table}

Despite the possibility to use human actors to determine features and values to convey emotions or the findings done in the works previously discussed, there is a need to get a better understanding on how convey emotions with non-anthropomorphic platform. This is due to the following reasons. First, it is not possible to apply a direct mapping from human studies to robots~\cite{Saerbeck2007,Canamero2010} due to robots' physical capabilities. Second, the significance of the agreement obtained from the studies that used human actors is on discussion~\cite{Russell2003}. Third, no all possible features present in a non-anthropomorphic platform are studied in the literature (e.g. angular velocity).
